---
title: "Qiime2.4_18S_HFP_2022"
author: "Sarah Tucker"
date: "12/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://github.com/benjjneb/dada2/issues/693

truncLen does two things: It truncates reads to the defined length AND it removes reads that don't reach the truncation length.

So using no truncation you lose no reads that way. When you trancate at say 145 nts you lose most of your reads that don't reach that far. And when you truncate to say 120 you keep most reads, since most reads reach that far.



This is the correct file! 




```{bash}
screen -S Miseq 

source /opt/anvio_conda/miniconda3/bin/activate bio-gen
conda activate qiime2-2021.4

qiime tools import --show-importable-types

cd /tank/tucker_data/HFP_18S/210216/210216_M01498_0767_000000000-JFNB9/BaseCalls



qiime tools import \
  --type 'SampleData[PairedEndSequencesWithQuality]' \
  --input-path /tank/tucker_data/HFP_18S/210216/210216_M01498_0767_000000000-JFNB9/BaseCalls\
  --input-format CasavaOneEightSingleLanePerSampleDirFmt \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/demux-paired-end.qza 
  
  
  
qiime demux summarize \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/demux-paired-end.qza    \
  --o-visualization  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/demux-paired-end.qzv


```


```{bash}
#1380F CCCTGCCHTTTGTACACAC
#1510R CCTTCYGCAGGTTCACCTAC
#1389F TTGTACACACCGCCC

#based on choi and beccas notes I suspect that 1880F and 1510R are what we are looking for 


qiime cutadapt trim-paired \
    --i-demultiplexed-sequences /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/demux-paired-end.qza \
    --p-adapter-f CCCTGCCHTTTGTACACAC...GTAGGTGAACCTGCRGAAGG \
    --p-adapter-r CCTTCYGCAGGTTCACCTAC...GTGTGTACAAADGGCAGGG \
    --o-trimmed-sequences /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/1380F-210216-demux-paired-end-trimmed.qza \
    --p-cores 30\
    --verbose \
    &> /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/1380F-210216-primer_trimming.log
    
    
    
qiime demux summarize \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/1380F-210216-demux-paired-end-trimmed.qza  \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/1380F-210216-demux-paired-end-trimmed.qzv

    
    

```


```{bash}

qiime dada2 denoise-paired \
  --i-demultiplexed-seqs /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/demux-paired-end.qza  \
  --p-trunc-len-f 125 \
  --p-trunc-len-r 125 \
  --o-representative-sequences /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-rep-seqs-dada2_AllSeqs.qza \
  --o-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-table-dada2_AllSeqs.qza \
  --o-denoising-stats /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-stats-dada2_AllSeqs.qza \
  --p-n-threads 30
  
  
qiime metadata tabulate \
  --m-input-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-stats-dada2_AllSeqs.qza  \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-stats-dada2_AllSeqs.qzv


  
qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-rep-seqs-dada2_AllSeqs.qza  \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-rep-seqs-dada2_AllSeqs.qzv


qiime feature-table summarize \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-table-dada2_AllSeqs.qza  \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-table-dada2_AllSeqs.qzv \
  --m-sample-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library1.tsv
  

  qiime feature-table filter-samples \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-table-dada2_AllSeqs.qza\
  --m-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library1.tsv \
  --p-where "[Library]='Library 1'" \
  --o-filtered-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-filtered-table-dada2_AllSeqs.qza 



qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-filtered-table-dada2_AllSeqs.qza    \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-exported-filtered-feature-table

cd /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-exported-filtered-feature-table
biom convert -i feature-table.biom -o 125-210216-exported-filtered-feature-table.tsv --to-tsv



qiime feature-table filter-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-rep-seqs-dada2_AllSeqs.qza  \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-filtered-table-dada2_AllSeqs.qza  \
  --o-filtered-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-filtered-rep-seqs.qza

qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-filtered-rep-seqs.qza \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-filtered-rep-seqs.qzv

  
  
  
  
qiime feature-table group \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-_210216-filtered-table-dada2_AllSeqs.qza \
  --p-axis sample \
  --m-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library1.tsv  \
  --m-metadata-column good_name \
  --p-mode sum \
  --o-grouped-table /tank/tucker_data/MiSeq_Analysis_2021_Final/Renamed-125-210216-filtered-table-dada2_AllSeqs.qza

cp /tank/tucker_data/MiSeq_Analysis_2021_Final/Renamed-125-210216-filtered-table-dada2_AllSeqs.qza  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125-210216-filtered-table-dada2_AllSeqs.qza

qiime feature-table summarize \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125-210216-filtered-table-dada2_AllSeqs.qza \
  --o-visualization  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-251-2021-single_table-dada2_AllSeqs.qzv \
  --m-sample-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed_Meta_seq_enviro_Library1.tsv



```





################
################

```{bash}
qiime tools import \
  --type 'SampleData[PairedEndSequencesWithQuality]' \
  --input-path /tank/tucker_data/HFP_18S/210219 \
  --input-format CasavaOneEightSingleLanePerSampleDirFmt \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Analysis_210219/demux-paired-end.qza 


###note in this round I did NOT Trim primers


```




```{bash}
qiime dada2 denoise-paired \
  --i-demultiplexed-seqs  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Analysis_210219/demux-paired-end.qza  \
  --p-trunc-len-f 125 \
  --p-trunc-len-r 125 \
  --o-representative-sequences /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-rep-seqs-dada2_AllSeqs.qza \
  --o-table  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-table-dada2_AllSeqs.qza \
  --o-denoising-stats  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-stats-dada2_AllSeqs.qza \
  --p-n-threads 30
  
  
  
  
qiime metadata tabulate \
  --m-input-file  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-stats-dada2_AllSeqs.qza \
  --o-visualization  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-stats-dada2_AllSeqs.qzv


  
qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-rep-seqs-dada2_AllSeqs.qza  \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-rep-seqs-dada2_AllSeqs.qzv


############
qiime feature-table summarize \
  --i-table  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-table-dada2_AllSeqs.qza  \
  --o-visualization  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-table-dada2_AllSeqs.qzv \
  --m-sample-metadata-file  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library2.tsv
  


  
  qiime feature-table filter-samples \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-table-dada2_AllSeqs.qza\
  --m-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library2.tsv \
  --p-where "[Library]='Library 2'" \
  --o-filtered-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-table-dada2_AllSeqs.qza 



qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-table-dada2_AllSeqs.qza    \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-exported-filtered-feature-table

cd /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-exported-filtered-feature-table
biom convert -i feature-table.biom -o 125_210219-exported-filtered-feature-table.tsv --to-tsv



qiime feature-table filter-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-rep-seqs-dada2_AllSeqs.qza  \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-table-dada2_AllSeqs.qza  \
  --o-filtered-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-rep-seqs.qza

qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-rep-seqs.qza \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-rep-seqs.qzv

  
  
  
  
qiime feature-table group \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-table-dada2_AllSeqs.qza \
  --p-axis sample \
  --m-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Meta_seq_enviro_Library2.tsv  \
  --m-metadata-column good_name \
  --p-mode sum \
  --o-grouped-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125_210219-filtered-table-dada2_AllSeqs.qza


qiime feature-table summarize \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125_210219-filtered-table-dada2_AllSeqs.qza \
  --o-visualization  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125_210219-filtered-table-dada2_AllSeqs.qzv \
  --m-sample-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed_Meta_seq_enviro_Library2.tsv



  
  
  
```





```{bash}


#######Time to merge!! 
qiime feature-table merge \
  --i-tables  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125-210216-filtered-table-dada2_AllSeqs.qza \
  --i-tables /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed-125_210219-filtered-table-dada2_AllSeqs.qza  \
  --o-merged-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF_table.qza



qiime feature-table merge-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-210216-rep-seqs-dada2_AllSeqs.qza  \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125_210219-filtered-rep-seqs.qza  \
  --o-merged-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF-rep-seqs.qza



qiime feature-table summarize \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF_table.qza \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF_table.qzv \
  --m-sample-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed_Filter_Meta_seq_enviro.tsv


qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF-rep-seqs.qza \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF-rep-seqs.qzv


###https://forum.qiime2.org/t/merging-feature-tables-should-i-also-merge-taxonomy-or-does-it-need-to-be-re-run/15678/2




```




```{bash}

qiime feature-table filter-features \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF_table.qza \
  --p-min-frequency 10 \
  --o-filtered-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/freq-125-MiSeq_18S_merged_HPF_table.qza
  

qiime feature-table filter-features \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/freq-125-MiSeq_18S_merged_HPF_table.qza  \
  --p-min-samples 2 \
  --o-filtered-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-MiSeq_18S_merged_HPF_table.qza  
  
  
  
qiime feature-table filter-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-MiSeq_18S_merged_HPF-rep-seqs.qza \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-MiSeq_18S_merged_HPF_table.qza  \
  --o-filtered-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF-rep-seqs.qza

qiime feature-table tabulate-seqs \
  --i-data /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF-rep-seqs.qza \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF-rep-seqs.qzv

  
  qiime feature-table summarize \
  --i-table /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-MiSeq_18S_merged_HPF_table.qza   \
  --o-visualization /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-MiSeq_18S_merged_HPF_table.qzv  \
  --m-sample-metadata-file /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/Renamed_Filter_Meta_seq_enviro.tsv


```



###Note after filtering the merged data we are at 766 samples and down to 5,849 ASVs
###many ASVs found in 500+ samples, so looks like merging was okay 

```{bash}





 qiime tools export \
  --input-path  /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF-rep-seqs.qza \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125_exported-rep-seqs
  
  
  
#freq-filtered-125-dna-sequences.fasta


qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-MiSeq_18S_merged_HPF_table.qza \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-exported-filtered-feature-table

cd /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filter-freq-125-exported-filtered-feature-table
biom convert -i feature-table.biom -o filter-freq-125-exported-filtered-feature-table.tsv --to-tsv






#https://docs.qiime2.org/2021.4/tutorials/filtering/


qiime phylogeny align-to-tree-mafft-fasttree \
  --i-sequences /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF-rep-seqs.qza  \
  --o-alignment /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-aligned_MiSeq_18S_merged_HPF-rep-seqs.qza  \
  --o-masked-alignment /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-masked_aligned_MiSeq_18S_merged_HPF-rep-seqs.qza \
  --o-tree /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF_unrooted-tree.qza \
  --o-rooted-tree /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF_rooted-tree.qza



#Saved FeatureData[AlignedSequence] to: single_aligned-rep-seqs.qza
#Saved FeatureData[AlignedSequence] to: masked-single_aligned-rep-seqs.qza
#Saved Phylogeny[Unrooted] to: unrooted-tree.qza
#Saved Phylogeny[Rooted] to: rooted-tree.qza



##more prep
qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF_unrooted-tree.qza  \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-unrooted-exported-tree



qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-MiSeq_18S_merged_HPF_rooted-tree.qza \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-rooted-exported-tree

#qiime tools export \
#  --input-path /tank/tucker_data/MiSeq_Analysis_2021_Final/MiSeq_merged_2021_2019_taxonomy.qza \
#  --output-path  /tank/tucker_data/MiSeq_Analysis_2021_Final/exported-taxonomy
 
 
 
 
 qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-masked_aligned_MiSeq_18S_merged_HPF-rep-seqs.qza  \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-masked-aligned-exported-rep-seqs
 
  
 qiime tools export \
  --input-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/filtered-freq-125-aligned_MiSeq_18S_merged_HPF-rep-seqs.qza \
  --output-path /tank/tucker_data/HFP_18S/Qiime_Analysis18S_2021/125-aligned-exported-rep-seqs


```





```{r}
#if (!requireNamespace("BiocManager", quietly=TRUE))
#   install.packages("BiocManager")
#BiocManager::install("DECIPHER")


#install.packages(devtools)
#devtools::install_github("pr2database/pr2database")

library(DECIPHER)
library(pr2database)


# Read training set 
trainingSet <- readRDS("/Users/sarahtucker/Documents/KByt/Phyloseq_2019_2021_MiSeq/pr2_version_4.14.0_SSU.decipher.trained.rds")

# Read sequences to assign
seq <- readDNAStringSet("/Users/sarahtucker/Documents/KByt/Analysis_18S/freq-filtered-125-dna-sequences.fasta")

# Get the taxonomy from the training set

  ids <- IdTaxa(seq,
                trainingSet,
                type="extended",
                strand="top",
                threshold=60)
#ids threshold is 60 now, before did not set a threshold
  
  #n_seq <- length(ids)
  #df_rows <- list()
taxo_levels <- c("kingdom", "supergroup", "division", "class", "order", "family", "genus", "species")

  
#for(i in 1:n_seq){
 # seq_name <- names(ids[i])
 # taxonomy=paste(x$taxon,collapse=";")
 # confidence <- ids[[i]]$confidence
 # df_rows[[i]] = data.frame(seq_name, taxonomy, confidence, taxo_level=c("Root", taxo_levels))
#}


#df <- reduce(df_rows, bind_rows) %>% 
 # filter(taxo_level !="Root") %>% 
 # pivot_wider(names_from = taxo_level, values_from = c(taxonomy, confidence))

###Code I used 
output <- sapply(ids,
function (id) {
seq_name <- names(ids)
paste(id$taxon,
";",
round(id$confidence, digits=1),
sep="",
collapse="; ")
})
head(output)

idk=as.data.frame(output)


write.csv(idk, "pr2_assigment_HFP_18S_125.csv")

pr2_18S_taxonomy=read.csv("pr2_assigment_HFP_18S_125_edited.csv")



```



#Becca removed contamination from water blanks and negative controls 
#https://benjjneb.github.io/decontam/vignettes/decontam_intro.html



```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("decontam")
```

```{r}
library(decontam)
# load libraries
library("ggplot2") 
library("phyloseq") 
#library(dada2)
library(methods)
library(plotly)
library(ggpubr)




```




```{r, echo=FALSE, warning=FALSE, message=FALSE}

getwd()

###now that you are in the right format you can bring things in as phyloseq objects

# read in otu table
otu_table_18SASV = read.csv("/Users/sarahtucker/Documents/KByt/Analysis_18S/edited_filter-freq-125-exported-filtered-feature-table.csv", sep=",", row.names=1) 
otu_table_18SASV = as.matrix(otu_table_18SASV)

# read in taxonomy
# seperated by kingdom phylum class order family genus species 

#you also need to edit your taxonomy table by separating each column by a comma so that you don't have just one column with all the taxonomic levels. Relabel to be  "Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species" as the column headers 

#note this was the original taxonomy, before using Pr2
taxonomy_18SASV = read.csv("/Users/sarahtucker/Documents/KByt/Analysis_18S/pr2_assigment_HFP_18S_125_edited.csv", sep=",", row.names=1) 


taxonomy_18SASV = as.matrix(taxonomy_18SASV)

#poop=subset(taxonomy_KBytNERR, Order=="D_3__Chloroplast")

#read in metadata
##note that there is not seq data for all metadata samples 
###when analyzing metadata strictly, need to edit
metadata_18SASV=read.csv("/Users/sarahtucker/Documents/KByt/Analysis_18S/Renamed_Filter_Meta_seq_enviro.csv", header=TRUE, fill=TRUE)

rownames(metadata_18SASV) <- metadata_18SASV$SampleID
metadata_18SASV$SampleID
META_18SASV= sample_data(metadata_18SASV)
nsamples(META_18SASV)

#read in tree
phy_tree18SASV = read_tree("/Users/sarahtucker/Documents/KByt/Analysis_18S/125-unrooted-tree.nwk")

# import as phyloseq objects
OTU_18SASV = otu_table(otu_table_18SASV, taxa_are_rows = TRUE) 
TAX_18SASV = tax_table(taxonomy_18SASV)


physeq_18SASV = phyloseq(OTU_18SASV, TAX_18SASV, phy_tree18SASV)

physeq_18SASV=merge_phyloseq(physeq_18SASV, META_18SASV)
ntaxa(physeq_18SASV)
#5849


```


```{r}



df <- as.data.frame(sample_data(physeq_18SASV)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(physeq_18SASV)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point()

```

```{r}
#Identify Contaminants - Prevalence

#In this method, the prevalence (presence/absence across samples) of each sequence feature in true positive samples is compared to the prevalence in negative controls to identify contaminants.


sample_data(physeq_18SASV)$is.neg <- sample_data(physeq_18SASV)$Sample_or_Control == "Control Negative"
contamdf.prev <- isContaminant(physeq_18SASV, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev$contaminant)
head(which(contamdf.prev$contaminant))



#The frequency of each sequence (or OTU) in the input feature table as a function of the concentration of amplified DNA in each sample is used to identify contaminant sequences.

#Threshold(Optional). Default 0.1. The probability threshold below which (strictly less than) the null-hypothesis (not a contaminant) should be rejected in favor of the alternate hypothesis (contaminant). A length-two vector can be provided when using the either or both methods: the first value is the threshold for the frequency test and the second for the prevalence test.

#In the prevalence test there is a special value worth knowing, threshold=0.5, that will identify as contaminants all sequences thare are more prevalent in negative controls than in positive samples. Let’s try using this more aggressive classification threshold rather than the default.

??isContaminant()


# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(physeq_18SASV, function(abund) 1*(abund>0))


ps.pa.neg <- prune_samples(sample_data(ps.pa)$Sample_or_Control == "Control Negative", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample_or_Control == "True Sample", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")





### explore negative controls
# prevelance-contaminants
ggarrange(
  plot_bar(subset_samples(
    prune_taxa(taxa_names(physeq_18SASV) %in% rownames(contamdf.prev[contamdf.prev$contaminant,]), physeq_18SASV), 
    Timeseries=="Neg"), x="Desc_Sample_ID", fill="Supergroup",
    title = "Contaiminants (Removed)"),
  plot_bar(subset_samples(
    prune_taxa(taxa_names(physeq_18SASV) %in% rownames(contamdf.prev[!contamdf.prev$contaminant,]), physeq_18SASV),
    Timeseries=="Neg"), x="Desc_Sample_ID", fill="Supergroup",
    title = "Not Contaiminants (Remaining)")
)


ps.noncontam <- prune_taxa(!contamdf.freq$contaminant, physeq_18SASV)
ps.noncontam



plyr::count(data.frame(OTU=rownames(contamdf.freq),
                 Frequency=contamdf.freq[,"contaminant"],
                 Prevelance=contamdf.prev[,"contaminant"]), vars = c("Frequency","Prevelance"))


# remove contaminants
ps.noncontam = prune_taxa(!contamdf.prev$contaminant, physeq_18SASV)
ntaxa(ps.noncontam)

contam = prune_taxa(contamdf.prev$contaminant, physeq_18SASV)
ntaxa(contam)
tax_table(contam)
write.csv(tax_table(contam), "/Users/sarahtucker/Documents/KByt/Analysis_18S/contamination_tax_table.csv")
write.csv(otu_table(contam), "/Users/sarahtucker/Documents/KByt/Analysis_18S/contamination_otu_table.csv")

##okay so taxonomically speaking the contaminants are all unimportant taxa (metazoa, fungi)





```



```{r}


# remove negative controls
ps.noncontam = subset_samples(ps.noncontam, Timeseries!="Neg")
ps.noncontam



# visualize metazoa
colors = data.frame()
metazoa = tax_table(tax_glom(ps.noncontam, "Phylum")); metazoa = metazoa[order(metazoa[,"Phylum"]),]
for (i in 1:dim(metazoa)[1]) {
  if (metazoa[i,"Phylum"]=="Metazoa") {
    colors[i,1] = "red"
  } else {
    colors[i,1] = "grey"
  }
}

plot1 = plot_bar(subset_samples(tax_glom(ps.noncontam, "Phylum"), Timeseries=="HFP"), x="Date", fill="Phylum") + 
  scale_y_continuous(limits = c(0,600000)) +
  labs(x = "Read Count", 
       title = "HFP", 
       caption = "Total abundances for metazoa (red) and unicellular (grey) reads") +
  theme_linedraw() +
  theme(legend.position = "none",
        plot.caption = element_text(hjust = 1),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = colors[,1])

plot2 = plot_bar(subset_samples(tax_glom(ps.noncontam, "Phylum"), Timeseries=="KBT"), x="Date", fill="Phylum") + 
  scale_y_continuous(limits = c(0,600000)) +
  labs(x = "", 
       title = "KBT", 
       caption = "") +
  theme_linedraw() +
  theme(legend.position = "none",
        plot.caption = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = colors[,1])
ggarrange(plot1, plot2, widths = c(3,1))
```

```{r}
# remove Metazoa
ps.noncontam.noMetazoa = subset_taxa(ps.noncontam, Phylum!="Metazoa")
ps.noncontam.noMetazoa

###note this didn't work because I didn't realize that there was a space involved in some of the taxa tables, this was resolved, but need to remove again in the clean up of the data 

ps.noncontam.noMetazoa.NoChano=subset_samples(ps.noncontam.noMetazoa, Timeseries!="Choano")
nsamples(ps.noncontam.noMetazoa.NoChano)



```


```{r}

write.csv(otu_table(ps.noncontam.noMetazoa.NoChano), "otu_table_Timeseries_ASV_18S_nocontam.csv")

write.csv(tax_table(ps.noncontam.noMetazoa.NoChano), "tax_table_Timeseries_ASV_18S_nocontam.csv")

write.csv(sample_data(ps.noncontam.noMetazoa.NoChano), "sample_data_Timeseries_ASV_18S_nocontam.csv")





```



